{\huge Changes and trends in budburst and leaf flush across Europe and North America} \\A meta-analysis of local adaptation in spring phenology studies

### 2022-12-23 Docs/manuscript.Rnw
updated methods

\subsection{Climate Data Gathering \& Analyses}
  \begin{itemize}
  \item Simple climate matrix (MAT \& MSP/MAP) gathering
  \begin{itemize}
  \item ClimateNA by Dr. Tongli Wang
  \item Climate Information Tool by FAO
  \end{itemize}
  \item Daily climate data gathering: Gridded daily climate data downloaded for all provenances and gardens: We hope that coarse metrics such as latitude and MAT ultimately represent how similar the climates are between the provenances and gardens in times that matter for the events. If climates are very similar, then we would expect similar timings [add more here].
To this end, we estimated climate overlap in relevant months: For spring events, we considered overlap across March to May.
  \begin{itemize}
  \item North American daily climate data from Daymet R package
  \item Europe daily climate data (gridded: NetCDF-4) retrieved from E-OBS 
  \end{itemize}
  \end{itemize}

\subsubsection{Estimate climate overlap using daily climate data using Overlap R package}
\subsubsection{Calculate GDDs}

@@

\subsubsection{Corrected for DOY (taking the difference)}\todo{will put results in supplement}
\subsection{Mixed effects modelling}

>>>>

\subsection{Climate Data Gathering \& Analyses}
With information about provenance latitude, longitude, and elevation from original publications, we estimated the mean annual temperature (MAT) for each provenance using the Climate Information Tool by FAO (Food and Agriculture Organization of the United Nations, 2022) and ClimateWNA (Wang et al., 2012). 
\\
To assess climate overlap percentage (define this better), we extracted gridded daily climate data for 2011 to 2020 for all European and North American provenances and gardens from E-OBS and the built-in Daymet package in R respectively (Cornes et al., 2018; Thornton et al., 2020).

\subsubsection{Estimate climate overlap using daily climate data using Overlap R package}
We hope that coarse metrics such as latitude and MAT ultimately represent how similar the climates are between the provenances and gardens in times that matter for the events. If climates are very similar, then we would expect similar timings [add more here].
To this end, we estimated climate overlap in relevant months: For spring events, we considered overlap across March to May.
\\
Using the Overlap package in R, we calculated the percentage overlap of the daily temperature of each provenance in spring months (March to May) from 2011 to 2020 in relation to the daily temperature of their corresponding gardens.

\subsubsection{Calculate GDDs}

@@

hmmm need to better define this... 
On top of taking the difference between provenance latitude and garden latitude (maybe I could write this in subscript), we also use the dism() function in the Geosphere package in R to estimate the geospherical distance between locations (accounting for Earth curvature).\improvement{maybe there are different/better terms}

\subsubsection{Corrected for DOY (taking the difference)}\todo{will put results in supplement}
(need to better define this as well)

\subsection{Mixed effects modelling}
